Parallel Global Search Algorithm with Local Tuning\\ for Solving Mixed-Integer Global Optimization Problems

### 2021-03-16 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
Some answers for reviewers comments

The first series of experiments was carried out using a sequential version of the algorithm without using local tuning in order to compare it with other similar methods.
The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox. Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}. The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods. All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM. The experimental results show the superiority of the PMGAL method both in terms of the number of iterations and the running time.

\begin{table}[ht]
	\caption{Comparison of the effectiveness of the PMGAL and GA methods}
	\label{tab:1}
	\center
	\begin{tabular}{|c|c|c|c|c|}
		\hline
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{PMGAL} \\
		\cline{2-3} \cline{4-5} 
		 & $k$ & $t$ &  $k$ & $t$  \\
		\hline 
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & 	417 &	0.04 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & 	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & 	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & 	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 &  372	 &	0.017 \\
		\hline
	\end{tabular}
\end{table}

@@

In total, we generated and solved 100 9-dimensional mixed-integer problems of this type. In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
The accuracy of the search was equal to $\epsilon = 10^{-2}$. Computational experiments were carried out on the Lobachevsky supercomputer. The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.

@@

The last series of experiments was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process. Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes. 
The acceleration provided by the parallel algorithm is not ideal, but quite acceptable.

@@

This article presents the results related to the development of parallel algorithms for global optimization for solving multiextremal problems in which some of the parameters are continuous, and some are discrete. 
These problems can also include non-convex constraints, which are accounted for using a special index scheme.
We propose a parallel algorithm for solving the problems of this class. The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.

The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak. This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.

In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.

@@

\bibitem{Zilinskas2011}
R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas and A.~Grothey, \textit{Parallel branch and bound for global optimization with combination of Lipschitz bounds}. Optimization Methods \& Software \textbf{26}(3), 487--498 (2011).

@@

>>>>

The first series of experiments was carried out using a sequential version of the algorithm without using local tuning in order to compare it with other similar methods.
The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox. Table \ref{tab:1} shows the number of trials and time it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.  
%Russian
Использованные тестовые задачи содержали от 2 до 4 непрерывных и от 2 до 4 дискретных параметров.

The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods.

All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM. The experimental results show the superiority of the PMGAL method both in terms of the number of iterations and the running time.

\begin{table}[ht]
	\caption{Comparison of the effectiveness of the PMGAL and GA methods}
	\label{tab:1}
	\center
	\begin{tabular}{|c|c|c|c|c|}
		\hline
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{PMGAL} \\
		\cline{2-3} \cline{4-5} 
		 & Number of trials & Time, sec. &  Number of trials & Time, sec.  \\
		\hline 
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & 	417 &	0.04 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & 	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & 	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & 	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 &  372	 &	0.017 \\
		\hline
	\end{tabular}
\end{table}

@@

In total, we generated and solved 100 mixed-integer problems of this type with 5 discrete and 4 continuous parameters. In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
The accuracy of the search was equal to $\epsilon = 10^{-2}$. Computational experiments were carried out on the Lobachevsky supercomputer. The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.

@@

The last experiment was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
%Russian
Параллельный алгоритм был применен для решения той же серии из 100 problems with 5 discrete and 4 continuous parameters.

Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process. Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes. 
The acceleration provided by the parallel algorithm is not ideal, but quite acceptable.

@@

This article presents the results related to the development of parallel algorithms for global optimization for solving multiextremal problems in which some of the parameters are continuous, and some are discrete. 
These problems can also include non-convex constraints, which are accounted for using a special index scheme.

%Russian
Основным предположением, которое обеспечивает применимость предложенного алгоритма, является выполнение условия Липшица для функций задачи при варьировании непрерывных параметров для каждого фиксированного набора дискретных параметров.
Задачи оптимизации, в которых функции являются липшицевыми, относятся к классу вычислительно-трудоемких задач. Любой численный алгоритм решения такой задачи требует вычисления значений функции в узлах некоторой (равномерной или неравномерной) сетки во всей области допустимых изменений параметров.
Несмотря на то, что предположение липшицевости позволяет провести оценку оптимума на существенно неравномерной (т.е. более экономной по сравнению с равномерной) сетке в области поиска, масштаб решаемых задач здесь ограничен. За приемлемое время даже с использованием параллельного алгоритма можно решать задачи, в которых порядка 10 непрерывных и 10 дискретных параметров. 
Решение задач с сотнями переменных может быть основано на схеме вложенной оптимизации в предположении, что зависимость целевой функции от части переменных носит локальный характер \cite{Barkalov2020_1}.

We propose a parallel algorithm for solving the problems of this class. The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.
The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak. This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.
In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.

%Russian
Примером сложной прикладной задачи, к которой планируется применить разработанные алгоритмы, является выбор оптимальных параметров конструкции aircraft hot-air anti-icing system для самолетов российского производства. В математической модели anti-icing system наряду с параметрами, изменяющимися непрерывно (например, диаметр трубки, подводящей теплый воздух от двигателя в крыло; диаметр выпускных отверстий на данной трубке; расстояние между отверстиями; и т.п.), есть и дискретные параметры (например, число рядов выпускных отверстий, которое из технических ограничений может быть равно 1, 2 или 3). Пример подобной постановки задачи см. в \cite{Habashi}.   Использование предложенного в статье алгоритма позволит не проводить оптимизацию для каждого значения дискретного параметра по отдельности, а заранее отсечь неперспективные варианты в процессе поиска.

@@

\bibitem{Zilinskas2011}
R.~Paulavi\v{c}ius, J.~\v{Z}ilinskas and A.~Grothey, \textit{Parallel branch and bound for global optimization with combination of Lipschitz bounds}, Optimization Methods \& Software \textbf{26}(3), 487--498 (2011).

@@

\bibitem{Barkalov2020_1}
K.~Barkalov, I.~Lebedev, M.~Kocheganova, V.~Gergel, \textit{Combining local and global search in a parallel nested optimization scheme}, Communications in Computer and Information Science, \textbf{1263}, 100-112 (2020).

\bibitem{Habashi}
M.~P.~C. Pellissier, W.~G. Habashi, A.~Pueyo. \textit{Optimization via FENSAP-ICE of aircraft hot-air anti-icing systems}, Journal of Aircraft \textbf{48}(1), 265--276 (2011).
